---
title: "Primatine OTC CEA"
author: "Joseph & Kate"
date: '2022-06-14'
output: html_document
editor_options: 
  chunk_output_type: console
---
######################################################################################################
####### Markov Model for CEA of Primatene OTC compared to Symbicort PRN and HCP management #### 2022 #
######################################################################################################

Objective: Cost-effectiveness of over the counter epinephrine vs. budesonide-formoterol in individuals with asthma not managed by a healthcare provider 
Study Population: Uninsured adults with asthma in the US using inhaler 2x weekly 
Setting and Location: US 
Study Perspective: US Societal 
Comparators: OTC Inhaled bud-form vs Inhaled epinephrine vs no OTC option
Model cycle: 1 week 
Time horizon: lifetime
Discount rate: 3% for both costs and QALYs 
Currency: 2021 USD 
Baseline age: 23 years old

Outcomes:  
- Longevity effects: Asthma-related mortality following sever exacerbation, background mortality from US life tables 
- Health-related quality-of-life effects: Utility of non-exacerbation, disutility of severe exacerbation 
Costs: 
- Medical costs: background direct medical costs, severe exacerbation medical costs
- Treatment costs: unit cost of Bud-form, Epi OTC (Primatine mist) 
- Time costs: Patient and caregiver time costs 
- Productivty loss: Presenteeism, absenteeism, unpaid work 

Assumptions 
1. No health insurance and inhaler use twice weekly (2x per week) 
2. Severe exacerbations require ED visit or hospitalization 
3. Patients who received ED care or Hospitalization were assumed to have been referred for care by HCP 
4. Patients under HCP management received daily ICS and as need albuterol (SABA) 
5. Risk reductions in severe exacerbations and asthma-related mortality were modeled for patients receiving healthcare provider management and PRN budesonide-formoterol OTC, with risk increases in asthma fatality modeled for patients receiving PRN inhaled epinephrine OTC 
6. Variable adherence to HCP management 


##################################### Model input #########################################

### Library ###
```{r,  Library, include=false}
library(here)
library(tidyverse)
library(dplyr)
library(readxl)
library(dampack)
library(ggplot2)
library(kableExtra)
library(purrr)

```

### Parameter Sheets ###
```{r, f_gen_base input}
f_gen_base <- function(n_sim = 1){
  
  require(truncnorm)  
  
### Asthma Control Probability ### (proportion: mean % of weeks with well controlled asthma)l
  p.H0_AC <- p.PRIM_AC <-  0.311
  p.SYM_AC <- 0.344
  p.H1_AC <- 0.444
  
### OTHER ### 
  #H0 entering H1
  p.H0_H1 <- 0
  
  #H1 Adherence
  pAnn.H1_H0  <- (1-0.79)                             # probability of non-adherence to HCP management
  p.H1_H0  <- 1 - (1-pAnn.H1_H0)^(1/52)

  #SEVR -> H1 (Recovery) - mean of 4 weeks
  wk.SEVR_H1 <- 4
  r.SEVR_H1 <- 1/wk.SEVR_H1                   #weekly rate of 1 event (event being leaving SEVR state)
  p.SEVR_H1 <- 1-exp(-r.SEVR_H1)
  
### SEV (Severe Exacerbation) ###
  p.SCS_H0 <- p.ED_H0 <-p.Hosp_H0 <- 0 #Assumption that following severe exacerbation patient enters H1
  RR.NC <- 1.201 #RATE ratio of exacerbation for not controlled asthma
  
  #Exacerbation proportion (SCS, ED, HOSP)
  prop.H0_SEV <- prop.PRIM_SEV <- c(0.776,0.130,0.094)
  prop.SYM_SEV <- c(0.819,0.102,0.079)
  prop.H1_SEV <- c(0.824,0.098,0.078)
  
  #Cumulative rate from SYGMA trials
  r.H0_SEVtot <- 0.20 #syg1
  r.SYM_SEVtot <- 0.11 #syg2
  r.H1_SEVtot <- 0.09 #syg1
  
    #Rate of exacerbation by control group
    #controlled (C)
    r.H0C_SEV <- r.H0_SEVtot/(p.H0_AC + RR.NC - RR.NC*p.H0_AC) #yaghoubi et al.2020
    r.SYMC_SEV <- r.SYM_SEVtot/(p.SYM_AC + RR.NC - RR.NC*p.SYM_AC) #yaghoubi et al.2020
    r.H1C_SEV <- r.H1_SEVtot/(p.H1_AC + RR.NC - RR.NC*p.H1_AC) #yaghoubi et al.2020
    
    #Not controlled (NC)
    r.H0NC_SEV <- r.H0C_SEV * RR.NC
    r.SYMNC_SEV <- r.SYMC_SEV * RR.NC
    r.H1NC_SEV <- r.H1C_SEV * RR.NC

 ## Transition probabilities for Severe exacerbation #
      #Severe Exacerbation Rates * Exacerbation Proportion
  
      #controlled (C)
        p.H0C_SCS <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[1] # weekly probability of severe exacerbation with no OTC or primatene OTC prn
        p.H0C_ED <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[2]
        p.H0C_Hosp <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[3]
        
        p.SYMC_SCS <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[1]	# weekly rate of severe exacerbation with symbicort PRN
        p.SYMC_ED <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[2]
        p.SYMC_Hosp <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[3]
        
        p.H1C_SCS <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[1]	# weekly rate of severe exacerbation with HCP management (low dose ICS + SABA)
        p.H1C_ED <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[2]
        p.H1C_Hosp <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[3]
      
      #Not controlled (NC)
        p.H0NC_SCS <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[1] # weekly probability of severe exacerbation with no OTC or primatene OTC prn
        p.H0NC_ED <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[2]
        p.H0NC_Hosp <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[3]
        
        p.SYMNC_SCS <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[1]	# weekly rate of severe exacerbation with symbicort PRN
        p.SYMNC_ED <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[2]
        p.SYMNC_Hosp <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[3]
        
        p.H1NC_SCS <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[1]	# weekly rate of severe exacerbation with HCP management (low dose ICS + SABA)
        p.H1NC_ED <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[2]
        p.H1NC_Hosp <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[3]
        
### Mortality parameters built in markov model ###
        
### Cost and utility inputs ###
  ## Direct Medical costs (per week)
  c.HCP     <- 53               # cost of HCP management
  c.AU      <- 15              # cost of uncontrolled asthma week
  c.SCS     <- 163               # cost of being in SCS 
  c.ED      <- 744               # cost of being in ED
  c.Hosp    <- 12398               # cost of being in HOSP
  
  ## Drug acquisition cost (per DAY) * 7 
  c.PRIM     <- 29.78*(0.52/160) * 7               # cost of primatene
  c.SYM      <- 311.49*(0.52/120) * 7            # cost of symbicort
  c.ICS      <- 326.28*(2/120) * 7               # cost of ICS
  c.SABA     <- 54.42*(0.49/200) * 7             # cost of SABA
  c.ICS_SABA <- c.ICS + c.SABA
  
  ## Indirect costs 
  c.timeHCP  <- 3.03 # HCP management time/week*Assuming 240 total minutes of 4 PCP visits per year, 20 min travel time, 40min wait, median wage $1037 in 2022
  c.prodAU   <- 132  # cost of productivity loss due to asthma not controlled days per week*
  
  #Productivity loss due to severe exacerbation ** Assuming do not go to work for 2,4,7 days
  c.prodSCS  <- 296 # productive lose due to severe exacerbation.
  c.prodED <- 593
  c.prodHosp <- 1037
  
  ## Utility
  du.AU   <-  -0.07/52                # disutility of asthma uncontrolled week
  du.SCS  <-  -0.10/52               # disutility of systemic corticosteroid use
  du.ED   <-  -0.15/52                 # disutility of emergency department
  du.Hosp <-  -0.20/52                 # disutility of hospitalization
  
  u.H0 <- 0.87/52          # utility of non-exacerbation
  u.H1 <- u.H0
  u.SCS <- u.H0 + du.SCS     # utility of SCS
  u.ED  <- u.H0 + du.ED        # utility of emergency deparment 
  u.Hosp<- u.H0 + du.Hosp   # utility of hospitalization

################################################################################################################################
### Parameter dataframe ###
df_psa <- data.frame(
    ## Asthma Control Probability (proportion: mean % of weeks with well controlled asthma)l
      p.H0_AC,
      p.PRIM_AC,
      p.SYM_AC,
      p.H1_AC,
      
    ## Transition probabilities #
      #controlled (C)
      p.H0C_SCS,
      p.H0C_ED,
      p.H0C_Hosp,
      
      p.SYMC_SCS,
      p.SYMC_ED,
      p.SYMC_Hosp,
      
      p.H1C_SCS,
      p.H1C_ED,
      p.H1C_Hosp,
      
      #Not controlled (NC)
      p.H0NC_SCS,
      p.H0NC_ED,
      p.H0NC_Hosp,
      
      p.SYMNC_SCS,
      p.SYMNC_ED,
      p.SYMNC_Hosp,
      
      p.H1NC_SCS,
      p.H1NC_ED,
      p.H1NC_Hosp,
    
      #SEVR -> H1
      p.SEVR_H1, #Weekly probability of exiting SEVR
    
      #H0
      p.H0_H1,
    
      #H1
      p.H1_H0,
    
    ### Cost and utility inputs ###
      ## Direct Medical costs (per week)
      c.HCP, 
      c.AU,
      c.SCS,
      c.ED,
      c.Hosp,
      
      ## Drug acquisition cost (per DAY) * 7 
      c.PRIM,
      c.SYM,
      c.ICS,
      c.SABA,
      c.ICS_SABA,
      
      ## Indirect costs 
      c.timeHCP,
      c.prodAU,
      
      #Productivity loss due to severe exacerbation ** Assuming do not go to work for 2,4,7 days
      c.prodSCS,
      c.prodED,
      c.prodHosp,
      
      ## Utility
      du.AU,
      du.SCS,
      du.ED,
      du.Hosp,
      
      u.H0,
      u.H1,
      u.SCS,
      u.ED,
      u.Hosp
      
  )
  
  v.du_exac  <- c(du.SCS, du.ED, du.Hosp)
  v.u_exac  <- c(u.SCS, u.ED, u.Hosp)
  return(df_psa)

}
```
```{r, f_gen_dsa input}
f_gen_dsa <- function(n_sim = 2,
                      DSA = FALSE){
  
  require(truncnorm)  

  r <- c(0.8,1.2)
### Asthma Control Probability ### (proportion: mean % of weeks with well controlled asthma)l
  p.H0_AC <- p.PRIM_AC <-  0.311
  p.SYM_AC <- 0.344
  p.H1_AC <- 0.444
  
### OTHER ###  
  #H0 entering H1
  pAnn.H0_H1 <- 0
  if(DSA=="H0_H1"){pAnn.H0_H1 <- c(0,0.2)}
  p.H0_H1  <- 1 - (1-pAnn.H0_H1)^(1/52)
  
  #H1 Adherence
  pAnn.H1_H0  <- (1-0.79)                             # probability of non-adherence to HCP management
  if(DSA=="H1_H0"){pAnn.H1_H0 <- 0.21 * r}
  p.H1_H0  <- 1 - (1-pAnn.H1_H0)^(1/52)

  #SEVR -> H1 (Recovery) - mean of 4 weeks
  wk.SEVR_H1 <- 4
  r.SEVR_H1 <- 1/wk.SEVR_H1                   #weekly rate of 1 event (event being leaving SEVR state)
  p.SEVR_H1 <- 1-exp(-r.SEVR_H1)
  
### SEV (Severe Exacerbation) ###
  p.SCS_H0 <- p.ED_H0 <-p.Hosp_H0 <- 0 #Assumption that following severe exacerbation patient enters H1
  RR.NC <- 1.201 #RATE ratio of exacerbation for not controlled asthma
  
  #Exacerbation proportion (SCS, ED, HOSP)
  prop.H0_SEV <- prop.PRIM_SEV <- c(0.776,0.130,0.094)
  prop.SYM_SEV <- c(0.819,0.102,0.079)
  prop.H1_SEV <- c(0.824,0.098,0.078)
  
  #Cumulative rate from SYGMA trials
  r.H0_SEVtot <- 0.20 #syg1
  if(DSA=="r.H0"){r.H0_SEVtot <- 0.20*r}
  
  r.SYM_SEVtot <- 0.11 #syg2
  if(DSA=="r.SYM"){r.SYM_SEVtot <- 0.11*r}

  r.H1_SEVtot <- 0.09 #syg1
  if(DSA=="r.H1"){r.H1_SEVtot <- 0.09*r}

    #Rate of exacerbation by control group
    #controlled (C)
    r.H0C_SEV <- r.H0_SEVtot/(p.H0_AC + RR.NC - RR.NC*p.H0_AC) #yaghoubi et al.2020
    r.SYMC_SEV <- r.SYM_SEVtot/(p.SYM_AC + RR.NC - RR.NC*p.SYM_AC) #yaghoubi et al.2020
    r.H1C_SEV <- r.H1_SEVtot/(p.H1_AC + RR.NC - RR.NC*p.H1_AC) #yaghoubi et al.2020
    
    #Not controlled (NC)
    r.H0NC_SEV <- r.H0C_SEV * RR.NC
    r.SYMNC_SEV <- r.SYMC_SEV * RR.NC
    r.H1NC_SEV <- r.H1C_SEV * RR.NC

 ## Transition probabilities for Severe exacerbation #
      #Severe Exacerbation Rates * Exacerbation Proportion
  
      #controlled (C)
        p.H0C_SCS <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[1] # weekly probability of severe exacerbation with no OTC or primatene OTC prn
        p.H0C_ED <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[2]
        p.H0C_Hosp <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[3]
        
        p.SYMC_SCS <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[1]	# weekly rate of severe exacerbation with symbicort PRN
        p.SYMC_ED <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[2]
        p.SYMC_Hosp <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[3]
        
        p.H1C_SCS <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[1]	# weekly rate of severe exacerbation with HCP management (low dose ICS + SABA)
        p.H1C_ED <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[2]
        p.H1C_Hosp <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[3]
      
      #Not controlled (NC)
        p.H0NC_SCS <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[1] # weekly probability of severe exacerbation with no OTC or primatene OTC prn
        p.H0NC_ED <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[2]
        p.H0NC_Hosp <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[3]
        
        p.SYMNC_SCS <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[1]	# weekly rate of severe exacerbation with symbicort PRN
        p.SYMNC_ED <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[2]
        p.SYMNC_Hosp <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[3]
        
        p.H1NC_SCS <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[1]	# weekly rate of severe exacerbation with HCP management (low dose ICS + SABA)
        p.H1NC_ED <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[2]
        p.H1NC_Hosp <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[3]
        
### Mortality parameters built in markov model ###
        
### Cost and utility inputs ###
  ## Direct Medical costs (per week)
  c.HCP     <- 53               # cost of HCP management
  if(DSA=="HCP"){c.HCP <- 53*r}
  
  c.AU      <- 15              # cost of uncontrolled asthma week
  if(DSA=="AU"){c.AU <- 15*r}
  
  c.SCS     <- 163               # cost of being in SCS 
  if(DSA=="SCS"){c.SCS <- 163*r}
  
  c.ED      <- 744               # cost of being in ED
  if(DSA=="ED"){c.ED <- 744*r}
  
  c.Hosp    <- 12398               # cost of being in HOSP
  if(DSA=="Hosp"){c.Hosp <- 12398*r}
  
  ## Drug acquisition cost (per DAY) * 7 
  #Med inhalations & Drug Acquisition
    inh.PRIM <- 0.52
    if(DSA=="inh.PRIM"){inh.PRIM <- 0.52*r}
    
    inh.SYM <- 0.52
    if(DSA=="inh.SYM"){inh.SYM <- 0.52*r}
    
    inh.ICS <- 2
    if(DSA=="inh.ICS"){inh.ICS <- 2*r}
    
    inh.SABA <- 0.49
    if(DSA=="inh.SABA"){inh.SABA <- 0.49*r}

  
    c.PRIM     <- 29.78*(inh.PRIM/160) * 7               # cost of primatene
    c.SYM      <- 311.49*(inh.SYM/120) * 7            # cost of symbicort
    c.ICS      <- 326.28*(inh.ICS/120)* 7               # cost of ICS
    c.SABA     <- 54.42*(inh.SABA/200)* 7             # cost of SABA
    c.ICS_SABA <- c.ICS + c.SABA

  
  ## Indirect costs 
  c.timeHCP  <- 3.03 # HCP management time/week*Assuming 240 total minutes of 4 PCP visits per year, 20 min travel time, 40min wait, median wage $1037 in 2022
  c.prodAU   <- 132  # cost of productivity loss due to asthma not controlled days per week*
  
  #Productivity loss due to severe exacerbation ** Assuming do not go to work for 2,4,7 days
  c.prodSCS  <- 296 # productive lose due to severe exacerbation.
  c.prodED <- 593
  c.prodHosp <- 1037

  
  ## Utility
  du.AU   <-  -0.07/52                # disutility of asthma uncontrolled week
  if(DSA=="du.AU"){du.AU <- (-0.07/52) * r}
  
  du.SCS  <-  -0.10/52               # disutility of systemic corticosteroid use
  if(DSA=="du.SCS"){du.SCS <- (-0.10/52) * r}

  du.ED   <-  -0.15/52                 # disutility of emergency department
  if(DSA=="du.ED"){du.ED <- (-0.15/52) * r}

  du.Hosp <-  -0.20/52                 # disutility of hospitalization
  if(DSA=="du.Hosp"){du.Hosp <- (-0.20/52) * r}

  
  u.H0 <- 0.87/52          # utility of non-exacerbation
  if(DSA=="u.H0"){u.H0 <- (0.87/52)*r}
  u.H1 <- u.H0
  u.SCS <- u.H0 + du.SCS     # utility of SCS
  u.ED  <- u.H0 + du.ED        # utility of emergency deparment 
  u.Hosp<- u.H0 + du.Hosp   # utility of hospitalization

################################################################################################################################
### Parameter dataframe ###
df_psa <- data.frame(
    ## Asthma Control Probability (proportion: mean % of weeks with well controlled asthma)l
      p.H0_AC,
      p.PRIM_AC,
      p.SYM_AC,
      p.H1_AC,
      
    ## Transition probabilities #
      #controlled (C)
      p.H0C_SCS,
      p.H0C_ED,
      p.H0C_Hosp,
      
      p.SYMC_SCS,
      p.SYMC_ED,
      p.SYMC_Hosp,
      
      p.H1C_SCS,
      p.H1C_ED,
      p.H1C_Hosp,
      
      #Not controlled (NC)
      p.H0NC_SCS,
      p.H0NC_ED,
      p.H0NC_Hosp,
      
      p.SYMNC_SCS,
      p.SYMNC_ED,
      p.SYMNC_Hosp,
      
      p.H1NC_SCS,
      p.H1NC_ED,
      p.H1NC_Hosp,
    
      #SEVR -> H1
      p.SEVR_H1, #Weekly probability of exiting SEVR
    
      #H0
      p.H0_H1,
    
      #H1
      p.H1_H0,
    
    ### Cost and utility inputs ###
      ## Direct Medical costs (per week)
      c.HCP, 
      c.AU,
      c.SCS,
      c.ED,
      c.Hosp,
      
      ## Drug acquisition cost (per DAY) * 7 
      c.PRIM,
      c.SYM,
      c.ICS,
      c.SABA,
      c.ICS_SABA,
      
      ## Indirect costs 
      c.timeHCP,
      c.prodAU,
      
      #Productivity loss due to severe exacerbation ** Assuming do not go to work for 2,4,7 days
      c.prodSCS,
      c.prodED,
      c.prodHosp,
      
      ## Utility
      du.AU,
      du.SCS,
      du.ED,
      du.Hosp,
      
      u.H0,
      u.H1,
      u.SCS,
      u.ED,
      u.Hosp
      
  )
  
  v.du_exac  <- c(du.SCS, du.ED, du.Hosp)
  v.u_exac  <- c(u.SCS, u.ED, u.Hosp)
  return(df_psa)

}

dsa_params<-c("H0_H1","H1_H0",
               "r.H0","r.SYM", "r.H1",
               "HCP","AU","SCS","ED","Hosp",
               "inh.PRIM","inh.SYM","inh.ICS","inh.SABA",
               "du.AU","du.SCS","du.ED","du.Hosp","u.H0")


```
```{r, f_gen_psa input}
psa_param <- read_excel("C:/Users/Joseph/Desktop/Grad studies MSc and Applications/AsthmaCEA2022/AsthmaCEA_Markov_2022_v6.xlsx", 
    sheet = "R-extract") %>% select(-"PSA distribution")
  t <- as_tibble(t(psa_param))
  names(t) <- t[1,]
  t<- t[-1,]
  t <- mutate_all(t, function(x) as.numeric(as.character(x)))
  rownames(t)<- c("S1", "S2")

psa_param <- apply(t,2,as.list)
rm(t)

f_gen_psa <- function(n_sim = 1000,
                      payer = "societal"){
  
  require(truncnorm)
  require(gtools)
  
### Asthma Control Probability ### (proportion: mean % of weeks with well controlled asthma)l
  p.H0_AC <- p.PRIM_AC <- rbeta(n = n_sim,
                                shape1 = psa_param[["p.H0_AC"]][["S1"]],
                                shape2 = psa_param[["p.H0_AC"]][["S2"]])
  p.SYM_AC <- rbeta(n = n_sim,
                    shape1 = psa_param[["p.SYM_AC"]][["S1"]],
                    shape2 = psa_param[["p.SYM_AC"]][["S2"]])
  
  p.H1_AC <- rbeta(n = n_sim,
                    shape1 = psa_param[["p.H1_AC"]][["S1"]],
                    shape2 = psa_param[["p.H1_AC"]][["S2"]])
  
### OTHER ###  
  #H0 entering H1
  p.H0_H1 <- 0
  
  #H1 Adherence
  pAnn.H1_H0  <- 1 - (rbeta(n = n_sim,
                    shape1 = psa_param[["p.H1_H0"]][["S1"]],
                    shape2 = psa_param[["p.H1_H0"]][["S2"]]))  # probability of non-adherence to HCP management

  p.H1_H0  <- 1 - (1-pAnn.H1_H0)^(1/52)
  
  
  #SEVR -> H1 (Recovery)
  wk.SEVR_H1 <- rexp(n_sim,1/4)      #mean of 4 weeks before exiting recovery state
  r.SEVR_H1 <- 1/wk.SEVR_H1      #weekly rate of 1 event (event being leaving SEVR state)
  p.SEVR_H1 <- 1-exp(-r.SEVR_H1)
  
  
### SEV (Severe Exacerbation) ###
  p.SCS_H0 <- p.ED_H0 <-p.Hosp_H0 <- 0 #Assumption that following severe exacerbation patient enters H1
  RR.NC <- rlnorm(n = n_sim,
                  meanlog = psa_param[["RR.NC"]][["S1"]],
                  sdlog = psa_param[["RR.NC"]][["S2"]]) #RATE ratio of exacerbation for not controlled asthma
  
  #Exacerbation proportion (SCS, ED, HOSP)
  prop.H0_SEV <- prop.PRIM_SEV <- rdirichlet(n = n_sim,
                                             c(psa_param[["prop.H0_SEV_SCS"]][["S1"]],psa_param[["prop.H0_SEV_ED"]][["S1"]],psa_param[["prop.H0_SEV_Hosp"]][["S1"]]))
  prop.SYM_SEV <- rdirichlet(n = n_sim,
                             c(psa_param[["prop.SYM_SEV_SCS2"]][["S1"]],psa_param[["prop.SYM_SEV_ED2"]][["S1"]],psa_param[["prop.SYM_SEV_Hosp2"]][["S1"]]))
  
  prop.H1_SEV <-  rdirichlet(n = n_sim,
                             c(psa_param[["prop.H1_SEV_SCS"]][["S1"]],psa_param[["prop.H1_SEV_ED"]][["S1"]],psa_param[["prop.H1_SEV_Hosp"]][["S1"]]))
  
  #Cumulative rate from SYGMA trials
  r.H0_SEVtot <- rgamma(n = n_sim,
                         shape = psa_param[["r.H0"]][["S1"]],
                         scale = psa_param[["r.H0"]][["S2"]])
                        
  r.SYM_SEVtot <- rgamma(n_sim,
                         shape = psa_param[["r.SYM_2"]][["S1"]],
                         scale = psa_param[["r.SYM_2"]][["S2"]])
  
  r.H1_SEVtot <-  rgamma(n_sim,
                         shape = psa_param[["r.H1"]][["S1"]],
                         scale = psa_param[["r.H1"]][["S2"]])
  
    #Rate of exacerbation by control group
    #controlled (C)
    r.H0C_SEV <- r.H0_SEVtot/(p.H0_AC + RR.NC - RR.NC*p.H0_AC) #yaghoubi et al.2020
    r.SYMC_SEV <- r.SYM_SEVtot/(p.SYM_AC + RR.NC - RR.NC*p.SYM_AC) #yaghoubi et al.2020
    r.H1C_SEV <- r.H1_SEVtot/(p.H1_AC + RR.NC - RR.NC*p.H1_AC) #yaghoubi et al.2020
    
    #Not controlled (NC)
    r.H0NC_SEV <- r.H0C_SEV * RR.NC
    r.SYMNC_SEV <- r.SYMC_SEV * RR.NC
    r.H1NC_SEV <- r.H1C_SEV * RR.NC

 #Rate -> transition probability
      p.H0C_SCS <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[,1] # weekly probability of severe exacerbation with no OTC or primatene OTC prn
      p.H0C_ED <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[,2]
      p.H0C_Hosp <- (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[,3]
      
      p.SYMC_SCS <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[,1]	# weekly rate of severe exacerbation with symbicort PRN
      p.SYMC_ED <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[,2]
      p.SYMC_Hosp <- (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[,3]
      
      p.H1C_SCS <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[,1]	# weekly rate of severe exacerbation with HCP management (low dose ICS + SABA)
      p.H1C_ED <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[,2]
      p.H1C_Hosp <- (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[,3]
      
      #Not controlled (NC)
      p.H0NC_SCS <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[,1] # weekly probability of severe exacerbation with no OTC or primatene OTC prn
      p.H0NC_ED <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[,2]
      p.H0NC_Hosp <- (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[,3]
      
      p.SYMNC_SCS <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[,1]	# weekly rate of severe exacerbation with symbicort PRN
      p.SYMNC_ED <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[,2]
      p.SYMNC_Hosp <- (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[,3]
      
      p.H1NC_SCS <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[,1]	# weekly rate of severe exacerbation with HCP management (low dose ICS + SABA)
      p.H1NC_ED <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[,2]
      p.H1NC_Hosp <- (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[,3]
    
    
### Mortality parameters built in markov model ### 
OR.AF <- rlnorm(n = n_sim,
                meanlog = psa_param[["OR.AF"]][["S1"]],
                sdlog = psa_param[["OR.AF"]][["S2"]]) #OR for saba overuse

#Cost
  #Med inhalations & Drug Acquisition
  inh.PRIM <- rgamma(n_sim,
                         shape = psa_param[["inh.PRIM"]][["S1"]],
                         scale = psa_param[["inh.PRIM"]][["S2"]])

  inh.SYM <- rgamma(n_sim,
                         shape = psa_param[["inh.SYM"]][["S1"]],
                         scale = psa_param[["inh.SYM"]][["S2"]])
  
  inh.ICS <- rgamma(n_sim,
                         shape = psa_param[["inh.ICS"]][["S1"]],
                         scale = psa_param[["inh.ICS"]][["S2"]])

  inh.SABA <- rgamma(n_sim,
                         shape = psa_param[["inh.SABA"]][["S1"]],
                         scale = psa_param[["inh.SABA"]][["S2"]])
  
      c.PRIM     <- 29.78*(inh.PRIM/160) * 7               # cost of primatene
      c.SYM      <- 311.49*(inh.SYM/120) * 7            # cost of symbicort
      c.ICS      <- 326.28*(inh.ICS/120)* 7               # cost of ICS
      c.SABA     <- 54.42*(inh.SABA/200)* 7             # cost of SABA
      c.ICS_SABA <- c.ICS + c.SABA
  
  #Direct Medical Costs
  c.HCP <- (rgamma(n_sim,
                        shape = psa_param[["c.HCP"]][["S1"]],
                        scale = psa_param[["c.HCP"]][["S2"]]))/52
  c.AU <- rgamma(n_sim,
                        shape = psa_param[["c.AU"]][["S1"]],
                        scale = psa_param[["c.AU"]][["S2"]])
  
  c.SCS <- rgamma(n_sim,
                        shape = psa_param[["c.SCS"]][["S1"]],
                        scale = psa_param[["c.SCS"]][["S2"]])
  c.ED <- rgamma(n_sim,
                        shape = psa_param[["c.ED"]][["S1"]],
                        scale = psa_param[["c.ED"]][["S2"]])
  c.Hosp <- rgamma(n_sim,
                        shape = psa_param[["c.Hosp"]][["S1"]],
                        scale = psa_param[["c.Hosp"]][["S2"]])
  
  #Indirect Costs
 if(payer=="societal"){ 
   
   c.timeHCP <- (rnorm(n = n_sim,
                     mean = psa_param[["c.timeHCP"]][["S1"]],
                     sd = psa_param[["c.timeHCP"]][["S2"]] )) * (1037/(60*52.1429*35))
  
  c.prodAU <- rlnorm(n = n_sim,
                  meanlog = psa_param[["c.prodAU"]][["S1"]],
                  sdlog = psa_param[["c.prodAU"]][["S2"]]) * 1037
  
  c.prodSCS <- rlnorm(n = n_sim,
                  meanlog = psa_param[["c.prodSCS"]][["S1"]],
                  sdlog = psa_param[["c.prodSCS"]][["S2"]]) * 1037
  
  c.prodED <- rlnorm(n = n_sim,
                  meanlog = psa_param[["c.prodED"]][["S1"]],
                  sdlog = psa_param[["c.prodED"]][["S2"]]) * 1037
  
  c.prodHosp <- rlnorm(n = n_sim,
                  meanlog = psa_param[["c.prodHosp"]][["S1"]],
                  sdlog = psa_param[["c.prodHosp"]][["S2"]]) * 1037
 }
  
if(payer=="public"){ 
 
  c.timeHCP <- 0
  c.prodAU <- 0
  c.prodSCS <- 0
  c.prodED <- 0
  c.prodHosp <- 0
 }
  
  #utility
  u.H0 <- (rbeta(n = n_sim,
                  shape1 = psa_param[["u.H0"]][["S1"]],
                  shape2 = psa_param[["u.H0"]][["S2"]]))/52
  
  du.AU <- -(rlnorm(n = n_sim,
                  meanlog = psa_param[["du.AU"]][["S1"]],
                  sdlog = psa_param[["du.AU"]][["S2"]]))/52
  
  du.SCS <- -(rbeta(n = n_sim,
                  shape1 = psa_param[["du.SCS"]][["S1"]],
                  shape2 = psa_param[["du.SCS"]][["S2"]]))/52
  du.ED <- -(rbeta(n = n_sim,
                  shape1 = psa_param[["du.ED"]][["S1"]],
                  shape2 = psa_param[["du.ED"]][["S2"]]))/52
  du.Hosp <- -(rbeta(n = n_sim,
                  shape1 = psa_param[["du.Hosp"]][["S1"]],
                  shape2 = psa_param[["du.Hosp"]][["S2"]]))/52
  
  u.SCS  <- (u.H0 + du.SCS)
  u.ED   <-  (u.H0 + du.ED)
  u.Hosp <- (u.H0 + du.Hosp) 
################################################################################################################################
### Parameter dataframe ###
df_psa <- data.frame(
    ## Asthma Control Probability (proportion: mean % of weeks with well controlled asthma)l
      p.H0_AC,
      p.PRIM_AC,
      p.SYM_AC,
      p.H1_AC,
      
    ## Transition probabilities #
      #Severe Exacerbation Rates * Exacerbation Proportion
      #controlled (C)
      p.H0C_SCS, 
      p.H0C_ED,
      p.H0C_Hosp,
      
      p.SYMC_SCS,
      p.SYMC_ED,
      p.SYMC_Hosp,
      
      p.H1C_SCS,
      p.H1C_ED,
      p.H1C_Hosp,
      
      #Not controlled (NC)
      p.H0NC_SCS,
      p.H0NC_ED,
      p.H0NC_Hosp,
      
      p.SYMNC_SCS,
      p.SYMNC_ED,
      p.SYMNC_Hosp,
      
      p.H1NC_SCS,
      p.H1NC_ED,
      p.H1NC_Hosp,
    
      #SEVR -> H1
      p.SEVR_H1, #Weekly probability of exiting SEVR
    
      #H0
      p.H0_H1,
    
      #H1
      p.H1_H0,               # probability of non-adherence to HCP management
    
        
    ### Cost and utility inputs ###
      ## Direct Medical costs (per week)
      c.HCP,
      c.AU,
      c.SCS,
      c.ED,
      c.Hosp,
      
      ## Drug acquisition cost (per DAY) * 7 
      c.PRIM,
      c.SYM,
      c.ICS,
      c.SABA,
      c.ICS_SABA,
      
      ## Indirect costs 
      c.timeHCP,
      c.prodAU,
      
      #Productivity loss due to severe exacerbation ** Assuming do not go to work for 2,4,7 days
      c.prodSCS,
      c.prodED,
      c.prodHosp,
      
      ## Utility
      du.AU,
      du.SCS,
      du.ED,
      du.Hosp,
      
      
      u.H0,          # utility of non-exacerbation
      u.H1 = u.H0, 
      u.SCS,     # utility of SCS
      u.ED,
      u.Hosp,
      OR.AF
      
  )
  
  v.du_exac  <- c(du.SCS, du.ED, du.Hosp)
  v.u_exac  <- c(u.SCS, u.ED, u.Hosp)
  return(df_psa)

}
```

### Markov Model ###
```{r, Markov Model, include=FALSE}
############################### Markov Model  ###########################
# The Markov function 
Markov <- function(params, v.M_1, n.t, v.n, d.c, d.e, trt=FALSE, res = TRUE) {
# Arguments:  
  # v.M_1:   initial allocation of cohort across states
  # n.t:     total number of cycles to run the model
  # v.n:     vector of health state names
  # d.c:     discount rate for costs
  # d.e:     discount rate for effectiveness (QALYs)
  # Trt:     is this the cohort receiving treatment? (default is FALSE)
  
  v.dwc <- 1 / (1 + d.c) ^ (0:n.t)   # calculate the cost discount weight based on the discount rate d.c 
  v.dwe <- 1 / (1 + d.e) ^ (0:n.t)   # calculate the QALY discount weight based on the discount rate d.e
  
  # create the transition trace matrix (m.TR) capturing the proportion of the cohort in each state at each time point
  m.TR <-  matrix(0, nrow = n.t + 1, ncol = n.s, 
                     dimnames = list( paste("cycle", 0:n.t, sep = ""), v.n))
  
  m.TR[1, ]  <- v.M_1  # indicate the initial health state
  
#Mortality: asthma fatality and baseline mortality (not affected by PSA)
  #Asthma Fatality (baseline)
    pAnn.SCS_D <- mortality$p.SCS_AF
    pAnn.ED_D <- mortality$p.ED_AF
    pAnn.Hosp_D <- mortality$p.HOSP_AF
    
    p.SCS_D <- 1 - (1-pAnn.SCS_D)^(1/52)
    p.ED_D <- 1 - (1-pAnn.ED_D)^(1/52)
    p.Hosp_D <- 1 - (1-pAnn.Hosp_D)^(1/52)
    
    #Asthma Fatality * Spitzer OR for SABA overuse
    pAnn.SCS_D_PRIM <- mortality$pSp.SCS_AF
    pAnn.ED_D_PRIM <- mortality$pSp.ED_AF
    pAnn.Hosp_D_PRIM <- mortality$pSp.HOSP_AF
    
    p.SCS_D_PRIM <- 1 - (1-pAnn.SCS_D_PRIM)^(1/52)
    p.ED_D_PRIM <- 1 - (1-pAnn.ED_D_PRIM)^(1/52)
    p.Hosp_D_PRIM <- 1 - (1-pAnn.Hosp_D_PRIM)^(1/52)
    
  #H0 Baseline Mortality
    pAnn.H0C_D <- pAnn.H0NC_D <- mortality$b_mortality   # probability of death from no HCP management
    p.H0C_D <- p.H0NC_D  <- p.H1C_D <- p.H1NC_D <-  1 - (1-pAnn.H0C_D)^(1/52)     

with(as.list(params),{  
   
  if(trt=="primatene"){
  
for (i in 2:(n.t + 1)) { 
    
  H0C.rem <- min((p.H0_H1*p.H1_AC+p.H0_H1*(1-p.H1_AC)+p.H0C_SCS+p.H0C_ED+p.H0C_Hosp+p.H0C_D[i-1]),1)
  H0NC.rem <- min((p.H0_H1*p.H1_AC+p.H0_H1*(1-p.H1_AC)+p.H0NC_SCS+p.H0NC_ED+p.H0NC_Hosp+p.H0NC_D[i-1]),1)
  H1C.rem <- min(((p.H1_H0*p.H0_AC)+(p.H1_H0*(1-p.H0_AC))+p.H1C_SCS+p.H1C_ED+p.H1C_Hosp+p.H1C_D[i-1]),1)
  H1NC.rem <- min(((p.H1_H0*p.H0_AC)+(p.H1_H0*(1-p.H0_AC))+p.H1NC_SCS+p.H1NC_ED+p.H1NC_Hosp+p.H1C_D[i-1]),1)
  
  SCS.rem <- min((p.SCS_D_PRIM[i-1]+p.H0C_D[i-1]),1)
  ED.rem <- min((p.ED_D_PRIM[i-1]+p.H0C_D[i-1]),1)
  Hosp.rem <-min((p.Hosp_D_PRIM[i-1]+p.H0C_D[i-1]),1)
  SEVR.rem <- min((p.SEVR_H1*p.H1_AC + p.SEVR_H1*(1-p.H1_AC)+p.H0C_D[i-1]),1)

    
  # create the transition matrix for each model cycle
    m.P <- matrix(c(
                    (1-H0C.rem)*p.H0_AC,(1-H0C.rem)*(1-p.H0_AC), p.H0_H1*p.H1_AC, p.H0_H1*(1-p.H1_AC), p.H0C_SCS, p.H0C_ED, p.H0C_Hosp, 0, p.H0C_D[i-1], #H0C
                    (1-H0NC.rem)*p.H0_AC,(1-H0NC.rem)*(1-p.H0_AC), p.H0_H1*p.H1_AC, p.H0_H1*(1-p.H1_AC), p.H0NC_SCS, p.H0NC_ED, p.H0NC_Hosp, 0, p.H0NC_D[i-1], #H0NC 
                    p.H1_H0*p.H0_AC,p.H1_H0*(1-p.H0_AC),(1 - H1C.rem)*p.H1_AC,(1 - H1C.rem)*(1-p.H1_AC),p.H1C_SCS,p.H1C_ED,p.H1C_Hosp, 0, p.H1C_D[i-1], #H1C
                    p.H1_H0*p.H0_AC,p.H1_H0*(1-p.H0_AC),(1 - H1NC.rem)*p.H1_AC,(1 - H1NC.rem)*(1-p.H1_AC),p.H1NC_SCS,p.H1NC_ED,p.H1NC_Hosp, 0, p.H1NC_D[i-1], #H1NC
                    0, 0, 0, 0, 0, 0, 0,(1-SCS.rem), min(p.SCS_D_PRIM[i-1]+p.H0C_D[i-1],1), #SCS
                    0, 0, 0, 0, 0, 0, 0,(1-ED.rem), min(p.ED_D_PRIM[i-1]+p.H0C_D[i-1],1), #ED
                    0, 0, 0, 0, 0, 0, 0,(1-Hosp.rem), min(p.Hosp_D_PRIM[i-1]+p.H0C_D[i-1],1), # Hosp
                    0, 0, p.SEVR_H1*p.H1_AC, p.SEVR_H1*(1-p.H1_AC), 0, 0, 0, 1-SEVR.rem, p.H0C_D[i-1], #SEVR
                    0, 0, 0, 0, 0, 0, 0, 0, 1), #D
                  nrow = n.s, ncol = n.s, byrow = T,
                  dimnames = list(v.n, v.n))
    
    # print(m.P) # show transition matrix for each model cycle to make sure p.H0_D is working - can remove after debugging
    
    # calculate the proportion of the cohort in each state at time t
    m.TR[i, ] <- t(m.TR[i - 1, ]) %*% m.P

}  # close the loop for the individuals 


  # create vectors of utility and costs for each state
  v.c <- c(
           c.PRIM, #H0C
           c.PRIM + c.AU + c.prodAU, #H0NC
           c.ICS_SABA + c.HCP + c.timeHCP, #H1C
           c.ICS_SABA + c.HCP + c.timeHCP + c.AU + c.prodAU, #H1NC
           c.SCS + c.prodSCS, #SCS
           c.ED + c.prodED, #ED
           c.Hosp + c.prodHosp, #HOSP
           c.ICS_SABA + c.HCP + c.timeHCP + c.AU + c.prodAU, #SEVR
           0) #D
  
  v.u <- c(
           u.H0, #H0C
           u.H0 + du.AU, #H0NC
           u.H1, #H1C
           u.H1 + du.AU, #H1NC
           u.SCS, #SCS
           u.ED, #ED
           u.Hosp, #HOSP
           u.H1 + du.AU,
           0) #D
  
  } 
  
  else{
  if(trt=="symbicort"){
    
  # run the simulation for symbicort
 for (i in 2:(n.t + 1)) { 
    
H0C.rem <- min((p.H0_H1*p.H1_AC+p.H0_H1*(1-p.H1_AC)+p.SYMC_SCS+p.SYMC_ED+p.SYMC_Hosp+p.H0C_D[i-1]),1)
H0NC.rem <- min((p.H0_H1*p.H1_AC+p.H0_H1*(1-p.H1_AC)+p.SYMNC_SCS+p.SYMNC_ED+p.SYMNC_Hosp+p.H0NC_D[i-1]),1)
H1C.rem <- min(((p.H1_H0*p.H0_AC)+(p.H1_H0*(1-p.H0_AC))+p.H1C_SCS+p.H1C_ED+p.H1C_Hosp+p.H1C_D[i-1]),1)
H1NC.rem <- min(((p.H1_H0*p.H0_AC)+(p.H1_H0*(1-p.H0_AC))+p.H1NC_SCS+p.H1NC_ED+p.H1NC_Hosp+p.H1C_D[i-1]),1)

SCS.rem <- min((p.SCS_D[i-1]+p.H0C_D[i-1]),1)
ED.rem <- min((p.ED_D[i-1]+p.H0C_D[i-1]),1)
Hosp.rem <- min((p.Hosp_D[i-1]+p.H0C_D[i-1]),1)
SEVR.rem <- min((p.SEVR_H1*p.H1_AC + p.SEVR_H1*(1-p.H1_AC)+p.H0C_D[i-1]),1)


  # create the transition matrix for each model cycle
    m.P <- matrix(c(
                    (1-H0C.rem)*p.SYM_AC,(1-H0C.rem)*(1-p.SYM_AC), p.H0_H1*p.H1_AC, p.H0_H1*(1-p.H1_AC), p.SYMC_SCS, p.SYMC_ED, p.SYMC_Hosp, 0, p.H0C_D[i-1], #H0C
                    (1-H0NC.rem)*p.SYM_AC,(1-H0NC.rem)*(1-p.SYM_AC), p.H0_H1*p.H1_AC, p.H0_H1*(1-p.H1_AC), p.SYMNC_SCS, p.SYMNC_ED, p.SYMNC_Hosp, 0, p.H0NC_D[i-1], #H0NC 
                    p.H1_H0*p.SYM_AC,p.H1_H0*(1-p.SYM_AC),(1 - H1C.rem)*p.H1_AC,(1 - H1C.rem)*(1-p.H1_AC),p.H1C_SCS,p.H1C_ED,p.H1C_Hosp, 0, p.H1C_D[i-1], #H1C
                    p.H1_H0*p.SYM_AC,p.H1_H0*(1-p.SYM_AC),(1 - H1NC.rem)*p.H1_AC,(1 - H1NC.rem)*(1-p.H1_AC),p.H1NC_SCS,p.H1NC_ED,p.H1NC_Hosp, 0, p.H1NC_D[i-1], #H1NC
                    0, 0, 0, 0, 0, 0, 0,(1-SCS.rem), min(p.SCS_D[i-1]+p.H0C_D[i-1],1), #SCS
                    0, 0, 0, 0, 0, 0, 0,(1-ED.rem), min(p.ED_D[i-1]+p.H0C_D[i-1],1), #ED
                    0, 0, 0, 0, 0, 0, 0,(1-Hosp.rem), min(p.Hosp_D[i-1]+p.H0C_D[i-1],1), # Hosp
                    0, 0, p.SEVR_H1*p.H1_AC, p.SEVR_H1*(1-p.H1_AC), 0, 0, 0, 1-SEVR.rem, p.H0C_D[i-1], #SEVR
                    0, 0, 0, 0, 0, 0, 0, 0, 1), #D
                  nrow = n.s, ncol = n.s, byrow = T,
                  dimnames = list(v.n, v.n))
    
    # print(m.P) # show transition matrix for each model cycle to make sure p.H0_D is working - can remove after debugging
    
    # calculate the proportion of the cohort in each state at time t
    m.TR[i, ] <- t(m.TR[i - 1, ]) %*% m.P
  
}  # close the loop for the individuals

  # create vectors of utility and costs for each state
  v.c <- c(
           c.SYM, #H0C
           c.SYM + c.AU + c.prodAU, #H0NC
           c.ICS_SABA + c.HCP + c.timeHCP, #H1C
           c.ICS_SABA + c.HCP + c.timeHCP + c.AU + c.prodAU, #H1NC
           c.SCS + c.prodSCS, #SCS
           c.ED + c.prodED, #ED
           c.Hosp + c.prodHosp, #HOSP
           c.ICS_SABA + c.HCP + c.timeHCP + c.AU + c.prodAU, #SEVR
           0) #D
  
  v.u <- c(
           u.H0, #H0C
           u.H0 + du.AU, #H0NC
           u.H1, #H1C
           u.H1 + du.AU, #H1NC
           u.SCS, #SCS
           u.ED, #ED
           u.Hosp, #HOSP
           u.H1 + du.AU, #SEVR
           0) #D
  
  }
    
  else{ #no treatment


# run the simulation for no treatment

for (i in 2:(n.t + 1)) { 
  
H0C.rem <- min((p.H0_H1*p.H1_AC+p.H0_H1*(1-p.H1_AC)+p.H0C_SCS+p.H0C_ED+p.H0C_Hosp+0+p.H0C_D[i-1]),1)
H0NC.rem <- min((p.H0_H1*p.H1_AC+p.H0_H1*(1-p.H1_AC)+p.H0NC_SCS+p.H0NC_ED+p.H0NC_Hosp+0+p.H0NC_D[i-1]),1)
H1C.rem <- min(((p.H1_H0*p.H0_AC)+(p.H1_H0*(1-p.H0_AC))+p.H1C_SCS+p.H1C_ED+p.H1C_Hosp+0+p.H1C_D[i-1]),1)
H1NC.rem <- min(((p.H1_H0*p.H0_AC)+(p.H1_H0*(1-p.H0_AC))+p.H1NC_SCS+p.H1NC_ED+p.H1NC_Hosp+0+p.H1C_D[i-1]),1)

SCS.rem <- min((p.SCS_D[i-1]+p.H0C_D[i-1]),1)
ED.rem <- min((p.ED_D[i-1]+p.H0C_D[i-1]),1)
Hosp.rem <- min((p.Hosp_D[i-1]+p.H0C_D[i-1]),1)
SEVR.rem <- min((p.SEVR_H1*p.H1_AC + p.SEVR_H1*(1-p.H1_AC)+p.H0C_D[i-1]),1)


  # create the transition matrix for each model cycle
    m.P <- matrix(c(
                    (1-H0C.rem)*p.H0_AC,(1-H0C.rem)*(1-p.H0_AC), p.H0_H1*p.H1_AC, p.H0_H1*(1-p.H1_AC), p.H0C_SCS, p.H0C_ED, p.H0C_Hosp, 0, p.H0C_D[i-1], #H0C
                    (1-H0NC.rem)*p.H0_AC,(1-H0NC.rem)*(1-p.H0_AC), p.H0_H1*p.H1_AC, p.H0_H1*(1-p.H1_AC), p.H0NC_SCS, p.H0NC_ED, p.H0NC_Hosp,0, p.H0NC_D[i-1], #H0NC 
                    p.H1_H0*p.H0_AC,p.H1_H0*(1-p.H0_AC),(1 - H1C.rem)*p.H1_AC,(1 - H1C.rem)*(1-p.H1_AC),p.H1C_SCS,p.H1C_ED,p.H1C_Hosp,0, p.H1C_D[i-1], #H1C
                    p.H1_H0*p.H0_AC,p.H1_H0*(1-p.H0_AC),(1 - H1NC.rem)*p.H1_AC,(1 - H1NC.rem)*(1-p.H1_AC),p.H1NC_SCS,p.H1NC_ED,p.H1NC_Hosp,0, p.H1NC_D[i-1], #H1NC
                    0, 0, 0, 0, 0, 0, 0,(1-SCS.rem), min(p.SCS_D[i-1]+p.H0C_D[i-1],1), #SCS
                    0, 0, 0, 0, 0, 0, 0,(1-ED.rem), min(p.ED_D[i-1]+p.H0C_D[i-1],1), #ED
                    0, 0, 0, 0, 0, 0, 0,(1-Hosp.rem), min(p.Hosp_D[i-1]+p.H0C_D[i-1],1), # Hosp
                    0, 0, p.SEVR_H1*p.H1_AC, p.SEVR_H1*(1-p.H1_AC), 0, 0, 0, 1-SEVR.rem, p.H0C_D[i-1], #SEVR
                    0, 0, 0, 0, 0, 0, 0, 0, 1), #D
                  nrow = n.s, ncol = n.s, byrow = T,
                  dimnames = list(v.n, v.n))
    
    # print(i-1)
    # print(p.H0C_D[i-1])
    m.TR[i, ] <- t(m.TR[i - 1, ]) %*% m.P
    # calculate the proportion of the cohort in each state at time t
  } # close the loop for the individuals 


  # create vectors of utility and costs for each state
  v.c <- c(
           0, #H0C
           c.AU + c.prodAU, #H0NC
           c.ICS_SABA + c.HCP + c.timeHCP, #H1C
           c.ICS_SABA + c.HCP + c.timeHCP + c.AU + c.prodAU, #H1NC
           c.SCS + c.prodSCS, #SCS
           c.ED + c.prodED, #ED
           c.Hosp + c.prodHosp, #HOSP
           c.ICS_SABA + c.HCP + c.timeHCP + c.AU + c.prodAU, #SEVR
           0) #D
  
  v.u <- c(
           u.H0, #H0C
           u.H0 + du.AU, #H0NC
           u.H1, #H1C
           u.H1 + du.AU, #H1NC
           u.SCS, #SCS
           u.ED, #ED
           u.Hosp, #HOSP
           u.H1 + du.AU, #SEVR
           0) #D
  

}
 
}

tc <-( m.TR %*% v.c)  # calculate the costs
te <-( m.TR %*% v.u ) # calculate the QALYs


tc_hat <- t(tc) %*% v.dwc  # total (discounted) cost 
te_hat <- t(te) %*% v.dwe  # total (discounted) QALY 

#outcomes
  m.TR_temp <- as_tibble(m.TR)
  
  cont <- sum(m.TR_temp$H0C + m.TR_temp$H1C)*7
  UC <- sum(m.TR_temp$H0NC + m.TR_temp$H1NC + m.TR_temp$SEVR)*7
  
  SCSoutcome <- sum(m.TR_temp$SCS)
  EDoutcome <- sum(m.TR_temp$ED)
  Hospoutcome <- sum(m.TR_temp$HOSP)
  
  SEVoutcome <- sum(m.TR_temp$SCS + m.TR_temp$ED + m.TR_temp$HOSP)

  if(res==TRUE){

results <- list(m.TR = m.TR, tc_hat = tc_hat, te_hat = te_hat, tc=tc, te=te, v.dwc=v.dwc, v.dwe=v.dwe, vc=v.c,
                controlled = cont, uncontrolled = UC, severe_exac = SEVoutcome) # store the results from the simulation in a list 


return(results)  # return the results
  } else{
    
     results <- c(
      "Cost" = tc_hat, 
      "QALY"   = te_hat, 
      "Control" = cont, 
      "Exac" = SEVoutcome
    )
    

return(results)  # return the results
  }
    
  
}) # end with function
}

```

### Wrapper Function (Markov + Parameter Sheet -> output) ###
```{r, Wrapper Function}
### Wrapper <- probablistic function
  f_wrapper <- function(fxn=f_gen_base(),
                        n_sim = 1000,
                        n.t = 52*77,
                        d.c = (0.03/52),
                        trt=FALSE){
    
    require(readr)
    require(tidyverse)

  ### Non-adjustable model inputs
    n.t   <- n.t             # time horizon, lifetime from 23yrs, 1 week cycle
    
    v.n   <- c("H0C","H0NC", #H0 no hcp mgt
               "H1C", "H1NC", #H1 HCP MGT
               "SCS","ED","HOSP", "SEVR", #SEV exacerbation
               "D") #death
    
    n.s   <- length(v.n)                 # the number of health states
    v.M_1 <- c(1,0,0,0,0,0,0,0,0)          # Initial health state distribution - Everyone starts at controlled no HCP management
    
    d.c   <- d.e <- d.c           # equal discounting of costs and QALYs by 3% per year
    v.Trt <- c("No OTC", 
               "Primatene", 
               "Symbicort")       # store the strategy names
  

  ### Reading in enviroment 
    environment(fxn) <- environment()
    environment(Markov)   <- environment()
    
#PSA parameters
    df_psa <- fxn
    
    #--  Run PSA  --#
    
    # Initialize matrix of results outcomes
    m_out <- matrix(NaN, 
                    nrow = n_sim, 
                    ncol = 4,
                    dimnames = list(1:n_sim,
                                    c("Costs", "QALYs",
                                      "Control", "Exac")))
    
    # run model for each row of PSA inputs
    for(t in 1:n_sim){
      mortality <- mort.list[[t]]
      
      # store results in row of results matrix
      m_out[t,] <- Markov(df_psa[t,], v.M_1, n.t, v.n, d.c, d.e, trt = trt, res=FALSE) 
    } # close model loop
    
    
    #-- Return results --#
    
    # convert matrix to dataframe (for plots)
    df_out <- as.data.frame(m_out) 
    
    # output the dataframe from the function  
    return(df_out) 
    
  } # end of function
```

### Analysis ###
```{r, Base Case}
m <- read_csv("mortality.csv")
mort.list <- list(m)

################################# Cost-effectiveness analysis #############################
sim_base_no_trt  <- f_wrapper(n_sim=1)
sim_base_PRIM    <- f_wrapper(n_sim=1, trt = "primatene") 
sim_base_SYMB    <- f_wrapper(n_sim=1, trt="symbicort")  



# create the cost-effectiveness table
table_markov <- rbind(sim_base_no_trt, sim_base_PRIM, sim_base_SYMB)
table_markov$Strategy <- c("No treatment","Epinephrine","ICS-LABA")
table_markov <- select(table_markov, Strategy, everything())

# calculate ICER
table_markov <- table_markov %>%
                    arrange(desc(QALYs)) %>% # no treatment is dominated by symbicort 
                    mutate(delta_Costs = c(Costs[1]-Costs[3],Costs[2]-Costs[3], NA),  # calculate incremental costs
                           delta_QALYs = c(QALYs[1]-QALYs[3],QALYs[2]-QALYs[3],NA),  # calculate incremental QALYs
                           delta_control = c(Control[1]-Control[3],Control[2]-Control[3],NA),
                           delta_exac = c(Exac[1]-Exac[3],Exac[2]-Exac[3],NA),
                           ICER=delta_Costs/delta_QALYs) # calculate the ICER

print(table_markov)
```
```{r, Probablistic}
### setting up PSA ###
  set.seed(1996) #reproducible results
  psa <- f_gen_psa(n_sim=1000) #PSA sheet
  mort.fxn <- function(){
        mortality <- read_csv("mortality.csv") %>% select(-pSp.SCS_AF,-pSp.ED_AF,-pSp.HOSP_AF)
    m <- rep(list(mortality),1000)
    ORF <- function(df, OR.AF = 2.6) {
        OR.AF <- OR.AF
        
        o.SCS  <- df$p.SCS_AF/(1-df$p.SCS_AF)
        o.ED   <- df$p.ED_AF/(1-df$p.ED_AF)
        o.HOSP <- df$p.HOSP_AF/(1-df$p.HOSP_AF)
        
        oSp.SCS  <- o.SCS*OR.AF
        oSp.ED   <-  o.ED *OR.AF
        oSp.HOSP <- o.HOSP*OR.AF
        
        pSp.SCS_AF  <- oSp.SCS/(1+oSp.SCS)
        pSp.ED_AF   <- oSp.ED/(1+oSp.ED)
        pSp.HOSP_AF <- oSp.HOSP/(1+oSp.HOSP)
        
        p <- cbind(df,pSp.SCS_AF,pSp.ED_AF,pSp.HOSP_AF)
        
      }
    t <- Map(ORF, m, OR.AF=psa$OR.AF)
  }
  mort.list <- mort.fxn()


################################# Cost-effectiveness analysis #############################
sim_PSA_no_trt  <- f_wrapper(fxn=psa, n_sim=1000)
sim_PSA_PRIM    <- f_wrapper(fxn=psa, n_sim=1000, trt = "primatene") 
sim_PSA_SYMB    <- f_wrapper(fxn=psa, n_sim=1000, trt="symbicort")  


#All results
  sim_PSA_no_trt$trt <- "noTrt"
  sim_PSA_PRIM$trt <- "PRIM"
  sim_PSA_SYMB$trt <- "SYMB"
  
  res <- rbind(sim_PSA_no_trt, sim_PSA_PRIM, sim_PSA_SYMB) 
  
  # write_csv(res,"C:\\Users\\Joseph\\Desktop\\Grad studies MSc and Applications\\AsthmaCEA2022\\Preliminary results\\PSAResults_ALLv4.csv")
  # write_csv(psa,"C:\\Users\\Joseph\\Desktop\\Grad studies MSc and Applications\\AsthmaCEA2022\\Preliminary results\\PSAParams_ALLv4.csv")


```
```{r, Scenario Analyses}
### setting up PSA ###
  set.seed(1996) #reproducible results
  psa <- f_gen_psa(n_sim=1000) #PSA sheet
  mort.fxn <- function(){
        mortality <- read_csv("mortality.csv") %>% select(-pSp.SCS_AF,-pSp.ED_AF,-pSp.HOSP_AF)
    m <- rep(list(mortality),1000)
    ORF <- function(df, OR.AF = 2.6) {
        OR.AF <- OR.AF
        
        o.SCS  <- df$p.SCS_AF/(1-df$p.SCS_AF)
        o.ED   <- df$p.ED_AF/(1-df$p.ED_AF)
        o.HOSP <- df$p.HOSP_AF/(1-df$p.HOSP_AF)
        
        oSp.SCS  <- o.SCS*OR.AF
        oSp.ED   <-  o.ED *OR.AF
        oSp.HOSP <- o.HOSP*OR.AF
        
        pSp.SCS_AF  <- oSp.SCS/(1+oSp.SCS)
        pSp.ED_AF   <- oSp.ED/(1+oSp.ED)
        pSp.HOSP_AF <- oSp.HOSP/(1+oSp.HOSP)
        
        p <- cbind(df,pSp.SCS_AF,pSp.ED_AF,pSp.HOSP_AF)
        
      }
    t <- Map(ORF, m, OR.AF=psa$OR.AF)
  }
  mort.list <- mort.fxn()

##Discounting##
  #0%
  sim_PSA_no_trt  <- f_wrapper(fxn=psa, d.c = 0, n_sim=1000)
  sim_PSA_PRIM    <- f_wrapper(fxn=psa, d.c = 0, n_sim=1000, trt = "primatene") 
  sim_PSA_SYMB    <- f_wrapper(fxn=psa, d.c = 0, n_sim=1000, trt="symbicort")  
  
    sim_PSA_no_trt$trt <- "noTrt"
    sim_PSA_PRIM$trt <- "PRIM"
    sim_PSA_SYMB$trt <- "SYMB"
    
    res_dc0 <- rbind(sim_PSA_no_trt, sim_PSA_PRIM, sim_PSA_SYMB) 

  #1.5%
  sim_PSA_no_trt  <- f_wrapper(fxn=psa, d.c = (0.015/52), n_sim=1000)
  sim_PSA_PRIM    <- f_wrapper(fxn=psa, d.c = (0.015/52), n_sim=1000, trt = "primatene") 
  sim_PSA_SYMB    <- f_wrapper(fxn=psa, d.c = (0.015/52), n_sim=1000, trt="symbicort")  
  
    sim_PSA_no_trt$trt <- "noTrt"
    sim_PSA_PRIM$trt <- "PRIM"
    sim_PSA_SYMB$trt <- "SYMB"
    
    res_dc1.5 <- rbind(sim_PSA_no_trt, sim_PSA_PRIM, sim_PSA_SYMB) 

##Time Horizon##
  # 10 years
  sim_PSA_no_trt  <- f_wrapper(fxn=psa, n.t = (10*52), n_sim=1000)
  sim_PSA_PRIM    <- f_wrapper(fxn=psa, n.t =  (10*52), n_sim=1000, trt = "primatene") 
  sim_PSA_SYMB    <- f_wrapper(fxn=psa, n.t =  (10*52), n_sim=1000, trt="symbicort")  
  
    sim_PSA_no_trt$trt <- "noTrt"
    sim_PSA_PRIM$trt <- "PRIM"
    sim_PSA_SYMB$trt <- "SYMB"
    
    res_th10 <- rbind(sim_PSA_no_trt, sim_PSA_PRIM, sim_PSA_SYMB)
    
  # 30 years
  sim_PSA_no_trt  <- f_wrapper(fxn=psa, n.t = (30*52), n_sim=1000)
  sim_PSA_PRIM    <- f_wrapper(fxn=psa, n.t =  (30*52), n_sim=1000, trt = "primatene") 
  sim_PSA_SYMB    <- f_wrapper(fxn=psa, n.t =  (30*52), n_sim=1000, trt="symbicort")  
  
    sim_PSA_no_trt$trt <- "noTrt"
    sim_PSA_PRIM$trt <- "PRIM"
    sim_PSA_SYMB$trt <- "SYMB"
    
    res_th30 <- rbind(sim_PSA_no_trt, sim_PSA_PRIM, sim_PSA_SYMB)
    
#Payer Perspective
  set.seed(1996) #reproducible results
  psa <- f_gen_psa(n_sim=1000, payer = "public") #PSA sheet
  
  sim_PSA_no_trt  <- f_wrapper(fxn=psa, n_sim=1000)
  sim_PSA_PRIM    <- f_wrapper(fxn=psa, n_sim=1000, trt = "primatene") 
  sim_PSA_SYMB    <- f_wrapper(fxn=psa, n_sim=1000, trt="symbicort")  
  
    sim_PSA_no_trt$trt <- "noTrt"
    sim_PSA_PRIM$trt <- "PRIM"
    sim_PSA_SYMB$trt <- "SYMB"
    
    res_public <- rbind(sim_PSA_no_trt, sim_PSA_PRIM, sim_PSA_SYMB)
    


```
```{r, Deterministic OSA}

#DSA setup
  dsa <- purrr::pmap(list(DSA=dsa_params),
                     f_gen_dsa) %>% bind_rows() #running all DSA parameters


#mortality setup
  m <- read_csv("mortality.csv")
  mort.list <- rep(list(m),nrow(dsa))

################################# Cost-effectiveness analysis #############################
sim_base_no_trt  <- f_wrapper(fxn=dsa, n_sim=nrow(dsa))
sim_base_PRIM    <- f_wrapper(fxn=dsa, n_sim=nrow(dsa), trt = "primatene") 
sim_base_SYMB    <- f_wrapper(fxn=dsa, n_sim=nrow(dsa), trt="symbicort")  

#naming function & setup for DSA
dsa_setup <- function(df,title){
  trt <- title
  
  dsa_params<-c("H0_H1","H1_H0",
               "r.H0","r.SYM", "r.H1",
               "HCP","AU","SCS","ED","Hosp",
               "inh.PRIM","inh.SYM","inh.ICS","inh.SABA",
               "du.AU","du.SCS","du.ED","du.Hosp","u.H0")
  
  ID <- 1:nrow(df)
  parameter <- rep(dsa_params,each=2)
  level <- rep(c("low","high"),nrow(dsa)/2)
  
  
  df <- cbind(ID,df,trt,parameter,level)

  }

  sim_base_no_trt  <- dsa_setup(sim_base_no_trt,"noTrt")
  sim_base_PRIM    <- dsa_setup(sim_base_PRIM, "PRIM")
  sim_base_SYMB    <- dsa_setup(sim_base_SYMB, "SYMB")

  res <- rbind(sim_base_no_trt,sim_base_PRIM,sim_base_SYMB)
  write_csv(res,"C:\\Users\\Joseph\\Desktop\\Grad studies MSc and Applications\\AsthmaCEA2022\\Preliminary results\\DSAResults.csv")
  
```

### Plotting ###
```{r, CEA plotting}
#Importing data
res <- read_csv("C:\\Users\\Joseph\\Desktop\\Grad studies MSc and Applications\\AsthmaCEA2022\\Preliminary results\\PSAResults_ALLv4.csv")
res$ID <- rep(1:1000,3)


#### Markov Table ####
  table_markov <- res %>% group_by(trt) %>% summarise(Costs = mean(Costs),
                                                      QALYs = mean(QALYs),
                                                      Control = mean(Control),
                                                      Exac = mean(Exac)
                                                      ) %>% mutate(delta_Costs = c(Costs[1]-Costs[2], NA, Costs[3]-Costs[2]),
                                                                   delta_QALYs = c(QALYs[1]-QALYs[2], NA, QALYs[3]-QALYs[2]), 
                                                                   delta_Control = c(Control[1]-Control[2], NA, Control[3]-Control[2]),
                                                                   delta_Exac = c(Exac[1]-Exac[2], NA, Exac[3]-Exac[2]),
                                                                   ICER=delta_Costs/delta_QALYs)

#results_wide
resw <- pivot_wider(res, names_from = trt, 
                        values_from = c(Costs, QALYs, Control, Exac))
  
  #Delta SYMB vs PRIM
  resw$deltaSP_Costs <- resw$Costs_SYMB - resw$Costs_PRIM
  resw$deltaSP_QALYs <- resw$QALYs_SYMB - resw$QALYs_PRIM
  
  #Delta noTrt vs PRIM
  resw$deltaNP_Costs <- resw$Costs_noTrt - resw$Costs_PRIM
  resw$deltaNP_QALYs <- resw$QALYs_noTrt - resw$QALYs_PRIM

### Cost-effectiveness Plane ###
  
  #single plot
    CEA_plot <- function(data, 
                         x, y, 
                         title=NULL,
                         xlab=NULL, ylab=NULL){
      p <- ggplot(data = data,
             aes(x,y))+
        geom_point()+
        stat_ellipse()+
        labs(x=xlab, y=ylab, title=title)+
        geom_hline(yintercept = 0)+
        geom_vline(xintercept = 0)+
        theme_bw() + theme(panel.border = element_blank(),
                           # panel.grid.major = element_blank(),
                           # panel.grid.minor = element_blank(), 
                           axis.line = element_line(colour = "black") 
                            )
        
    return(p)
    }
    
      CEA_plot(data = resw, x = resw$deltaSP_QALYs, y = resw$deltaSP_Costs, title = "SYMB")
      CEA_plot(data = resw, x = resw$deltaNP_QALYs, y = resw$deltaNP_Costs, title = "noTrt")
      
  #combined plot
     CEA_plot_JH <- function(res){
      
       #creating dataset
       res <- res %>% select(ID, trt, Costs, QALYs)
       
       SYMB <- res %>% filter(trt=="SYMB")
       noTrt <- res %>% filter(trt=="noTrt")
       PRIM <- res %>% filter(trt=="PRIM")
       
        SYMB$delta_Costs <- SYMB$Costs - PRIM$Costs
        SYMB$delta_QALYs <- SYMB$QALYs - PRIM$QALYs
        noTrt$delta_Costs <- noTrt$Costs - PRIM$Costs
        noTrt$delta_QALYs <- noTrt$QALYs - PRIM$QALYs
        
          res.temp <- rbind(SYMB, noTrt)

       
       #creating plot
       p <- ggplot(data = res.temp, aes(delta_QALYs,delta_Costs, group = trt, color = trt))+
        geom_point()+
        stat_ellipse(type = "norm", lty=2)+
        labs(x="QALYs", y="Costs")+
        geom_hline(yintercept = 0)+
        geom_vline(xintercept = 0)+
        theme_bw() + theme(panel.border = element_blank(),
                           # panel.grid.major = element_blank(),
                           # panel.grid.minor = element_blank(), 
                           axis.line = element_line(colour = "black"),
                           legend.title = element_blank(),
                           legend.position = "bottom"
                            ) 

       
    return(p)
     } 
     CEA_plot_JH(res)
  


#CEAC
  CEAC <- function(costs, qalys, title=NULL){
  
    WTP <- seq(from=0, to=150000, by=10000)


   #INMB matrix
    ceac.m <- data.frame(matrix(nrow=1000, ncol=length(WTP)))
    for(i in 1:length(WTP)){
    ceac.m[,i] <- qalys*WTP[i] - costs
    }
  
  #CEAC proportions
    ceac.prop <- data.frame(matrix(nrow=1, ncol=length(WTP)))
    for(i in 1:length(WTP)){
      ceac.prop[,i] <- sum(ceac.m[,i]>0) 
    }
  
      ceac.prop <- as_tibble(t(ceac.prop/1000)) %>% rename(prop = V1)
      ceac.prop$WTP <- WTP
    
  #Plotting 
     p <- ggplot(data = ceac.prop,
           aes(WTP,prop))+
      geom_line()+
      ylim(0,1)+
      labs(y="Proportion", x="Willingness-to-pay", title=title)+
      theme_bw() + theme(panel.border = element_blank(),
                         # panel.grid.major = element_blank(),
                         # panel.grid.minor = element_blank(), 
                         axis.line = element_line(colour = "black") 
                          )
      
      
  l<-list(ceac.m, ceac.prop, p)
  }

  SP <- CEAC(costs=resw$deltaSP_Costs, qalys=resw$deltaSP_QALYs) #SYMB vs PRIM
  NP <- CEAC(costs=resw$deltaNP_Costs, qalys=resw$deltaNP_QALYs) #No Trt vs PRIM

  p <- SP[[3]] +
    geom_line(aes(x=(NP[[2]][["WTP"]]), y=(NP[[2]][["prop"]]), color="red"))
  p
```
```{r, Tornado}
res <- read_csv("C:\\Users\\Joseph\\Desktop\\Grad studies MSc and Applications\\AsthmaCEA2022\\Preliminary results\\DSAResults.csv") 

#results_wide
resw <- pivot_wider(res, names_from = trt, 
                        values_from = c(Costs, QALYs, Control, Exac))
  
  #Delta SYMB vs PRIM
  resw$deltaSP_Costs <- resw$Costs_SYMB - resw$Costs_PRIM
  resw$deltaSP_QALYs <- resw$QALYs_SYMB - resw$QALYs_PRIM
  
  #Delta noTrt vs PRIM
  resw$deltaNP_Costs <- resw$Costs_noTrt - resw$Costs_PRIM
  resw$deltaNP_QALYs <- resw$QALYs_noTrt - resw$QALYs_PRIM

tornado_plot <- function(df, comparison=FALSE){
  
  resw <- resw %>% select(parameter,level,deltaSP_Costs,deltaSP_QALYs)
  resw$INMB <- resw$deltaSP_QALYs * 100000 - resw$deltaSP_Costs
  
  
  
  resw <- resw %>% select(parameter, level, INMB) %>%  pivot_wider(names_from = level, values_from = INMB) %>% rename(Parameter=parameter,
                                                                                                                      Lower_Bound = low,
                                                                                                                      Upper_Bound = high)
  df<-resw
  df$UL_Difference <- abs(df$Upper_Bound - df$Lower_Bound)
  
 
  
# get order of parameters according to size of intervals
order.parameters <- df %>% arrange(UL_Difference) %>%
  mutate(Parameter=factor(x=Parameter, levels=Parameter)) %>%
  select(Parameter) %>% unlist() %>% levels()

# width of columns in plot (value between 0 and 1)
width <- 0.95

# get data frame in shape for ggplot and geom_rect
df.2 <- df %>% 
  # gather columns Lower_Bound and Upper_Bound into a single column using gather
  gather(key='type', value='output.value', Lower_Bound:Upper_Bound) %>%
  # just reordering columns
  select(Parameter, type, output.value, UL_Difference) %>%
  # create the columns for geom_rect
  mutate(Parameter=factor(Parameter, levels=order.parameters),
         ymin=pmin(output.value, base.value),
         ymax=pmax(output.value, base.value),
         xmin=as.numeric(Parameter)-width/2,
         xmax=as.numeric(Parameter)+width/2)

ggplot() + 
  geom_rect(data = df.2, 
            aes(ymax=ymax, ymin=ymin, xmax=xmax, xmin=xmin, fill=type)) +
  theme_bw() + 
  theme(axis.title.y=element_blank(), legend.position = 'bottom',
        legend.title = element_blank()) + 
  geom_hline(yintercept = base.value) +
  scale_x_continuous(breaks = c(1:length(order.parameters)), 
                     labels = order.parameters) +
  coord_flip()
  
}

  




```

```{r, NMB for symbicort}
f_gen_psa_NMB <- function(n_sim = 1000){
  
  require(truncnorm)  
  
  ### Asthma Control Probability ### (proportion: mean % of weeks with well controlled asthma)l
  p.H0_AC <- p.PRIM_AC <-  0.31
  p.SYM_AC <- 0.34
  p.H1_AC <- 0.44
  
  ### OTHER ###  
  #H1 Adherence
  pAnn.H1_H0  <- (1-0.79)                             # probability of non-adherence to HCP management
  
  #SEVR -> H1 (Recovery)
  # wk.SEVR_H1 <- rexp(n_sim,1/4)      #mean of 4 weeks before exiting recovery state
  wk.SEVR_H1 <- 1/4
  r.SEVR_H1 <- 1/wk.SEVR_H1      #weekly rate of 1 event (event being leaving SEVR state)
  
  ### SEV (Severe Exacerbation) ###
  p.SCS_H0 <- p.ED_H0 <-p.Hosp_H0 <- 0 #Assumption that following severe exacerbation patient enters H1
  RR.NC <- 1.201 #RATE ratio of exacerbation for not controlled asthma
  
  #Exacerbation proportion (SCS, ED, HOSP)
  prop.H0_SEV <- prop.PRIM_SEV <- c(0.776,0.130,0.094)
  prop.SYM_SEV <- c(0.819,0.102,0.079)
  prop.H1_SEV <- c(0.784,0.152,0.064)
  
  #Cumulative rate from SYGMA trials
  r.H0_SEVtot <- 0.20 #syg1
  r.SYM_SEVtot <- 0.11 #syg2
  r.H1_SEVtot <- 0.09 #syg1
  
  #Rate of exacerbation by control group
  #controlled (C)
  r.H0C_SEV <- r.H0_SEVtot/(p.H0_AC + RR.NC - RR.NC*p.H0_AC) #yaghoubi et al.2020
  r.SYMC_SEV <- r.SYM_SEVtot/(p.SYM_AC + RR.NC - RR.NC*p.SYM_AC) #yaghoubi et al.2020
  r.H1C_SEV <- r.H1_SEVtot/(p.H1_AC + RR.NC - RR.NC*p.H1_AC) #yaghoubi et al.2020
  
  #Not controlled (NC)
  r.H0NC_SEV <- r.H0C_SEV * RR.NC
  r.SYMNC_SEV <- r.SYMC_SEV * RR.NC
  r.H1NC_SEV <- r.H1C_SEV * RR.NC
  
  ### Mortality ### 
  #Asthma Fatality (baseline)
  pAnn.SCS_D <- mortality$p.SCS_AF
  pAnn.ED_D <- mortality$p.ED_AF
  pAnn.Hosp_D <- mortality$p.HOSP_AF
  
  p.SCS_D <- 1 - (1-pAnn.SCS_D)^(1/52)
  p.ED_D <- 1 - (1-pAnn.ED_D)^(1/52)
  p.Hosp_D <- 1 - (1-pAnn.Hosp_D)^(1/52)
  
  #Asthma Fatality * Spitzer OR for SABA overuse
  pAnn.SCS_D_PRIM <- mortality$pSp.SCS_AF
  pAnn.ED_D_PRIM <- mortality$pSp.ED_AF
  pAnn.Hosp_D_PRIM <- mortality$pSp.HOSP_AF
  
  p.SCS_D_PRIM <- 1 - (1-pAnn.SCS_D_PRIM)^(1/52)
  p.ED_D_PRIM <- 1 - (1-pAnn.ED_D_PRIM)^(1/52)
  p.Hosp_D_PRIM <- 1 - (1-pAnn.Hosp_D_PRIM)^(1/52)
  
  #H0 Baseline Mortality
  pAnn.H0C_D <- pAnn.H0NC_D <- mortality$b_mortality   # probability of death from no HCP management
  p.H0C_D <- p.H0NC_D  <- p.H1C_D <- p.H1NC_D <-  1 - (1-pAnn.H0C_D)^(1/52)                 
  
  
  ################################################################################################################################
  ### Parameter dataframe ###
  df_psa <- data.frame(
    ## Asthma Control Probability (proportion: mean % of weeks with well controlled asthma)l
    p.H0_AC,
    p.PRIM_AC,
    p.SYM_AC,
    p.H1_AC,
    
    ## Transition probabilities #
    #Severe Exacerbation Rates * Exacerbation Proportion
    #controlled (C)
    p.H0C_SCS = (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[1], # weekly probability of severe exacerbation with no OTC or primatene OTC prn
    p.H0C_ED = (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[2],
    p.H0C_Hosp = (1-exp(-r.H0C_SEV/52)) * prop.H0_SEV[3],
    
    p.SYMC_SCS = (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[1],	# weekly rate of severe exacerbation with symbicort PRN
    p.SYMC_ED = (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[2],
    p.SYMC_Hosp = (1-exp(-r.SYMC_SEV/52)) * prop.SYM_SEV[3],
    
    p.H1C_SCS = (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[1],	# weekly rate of severe exacerbation with HCP management (low dose ICS + SABA)
    p.H1C_ED = (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[2],
    p.H1C_Hosp = (1-exp(-r.H1C_SEV/52)) * prop.H1_SEV[3],
    
    #Not controlled (NC)
    p.H0NC_SCS = (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[1], # weekly probability of severe exacerbation with no OTC or primatene OTC prn
    p.H0NC_ED = (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[2],
    p.H0NC_Hosp = (1-exp(-r.H0NC_SEV/52)) * prop.H0_SEV[3],
    
    p.SYMNC_SCS = (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[1],	# weekly rate of severe exacerbation with symbicort PRN
    p.SYMNC_ED = (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[2],
    p.SYMNC_Hosp = (1-exp(-r.SYMNC_SEV/52)) * prop.SYM_SEV[3],
    
    p.H1NC_SCS = (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[1],	# weekly rate of severe exacerbation with HCP management (low dose ICS + SABA)
    p.H1NC_ED = (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[2],
    p.H1NC_Hosp = (1-exp(-r.H1NC_SEV/52)) * prop.H1_SEV[3],
    
    #SEVR -> H1
    p.SEVR_H1 = 1-exp(-r.SEVR_H1), #Weekly probability of exiting SEVR
    
    #H0
    p.H0_H1 = 0,
    
    #H1
    p.H1_H0  = 1 - (1-pAnn.H1_H0)^(1/52),               # probability of non-adherence to HCP management
    
    
    ### Cost and utility inputs ###
    ## Direct Medical costs (per week)
    c.HCP     = 53,               # cost of HCP management
    c.AU      = 15,              # cost of uncontrolled asthma week
    c.SCS     = 163,               # cost of being in SCS 
    c.ED      = 744,               # cost of being in ED
    c.Hosp    = 12398,               # cost of being in HOSP
    
    ## Drug acquisition cost (per DAY) * 7 
    c.PRIM     = 29.78*(0.52/160) * 7,               # cost of primatene
    c.SYM      = (seq(from=0, to=500, length=n_sim))*(0.52/120) * 7,            # cost of symbicort
    c.ICS      =  c.ICS <- 326.28*(2/120)* 7,               # cost of ICS
    c.SABA     = c.SABA <- 54.42*(0.49/200)* 7,             # cost of SABA
    c.ICS_SABA = (c.ICS + c.SABA),
    
    ## Indirect costs 
    c.timeHCP  = 2, # HCP management time/week*Assuming 240 total minutes of 4 PCP visits per year, 20 min travel time, 40min wait, median wage $1037 in 2022
    c.prodAU   = 132, # cost of productivity loss due to asthma not controlled days per week*
    
    #Productivity loss due to severe exacerbation ** Assuming do not go to work for 2,4,7 days
    c.prodSCS  = 296, # productive lose due to severe exacerbation.
    c.prodED = 593,
    c.prodHosp = 1037,
    
    ## Utility
    du.AU   = du.AU <- -0.06/52,                # disutility of asthma uncontrolled week
    du.SCS  = du.SCS <- -0.10/52,               # disutility of systemic corticosteroid use
    du.ED   = du.ED <- -0.15/52,                 # disutility of emergency department
    du.Hosp = du.Hosp <- -0.20/52,                 # disutility of hospitalization
    
    
    u.H0 = u.H0 <- 0.87/52,          # utility of non-exacerbation
    u.H1 = u.H1 <- u.H0, 
    u.SCS = u.SCS <- (u.H0 + du.SCS),     # utility of SCS
    u.ED   = u.ED <- (u.H0 + du.ED),        # utility of emergency deparment 
    u.Hosp    = u.Hosp <- (u.H0 + du.Hosp)   # utility of hospitalization
    
  )
  
  v.du_exac  <- c(du.SCS, du.ED, du.Hosp)
  v.u_exac  <- c(u.SCS, u.ED, u.Hosp)
  return(df_psa)
  
}

NMB_plot <- function(data, 
                     x, y, 
                     title=NULL,
                     xlab=NULL, ylab=NULL){
  p <- ggplot(data = data,
         aes(x,y))+
    geom_line(lty=2)+
    labs(x=xlab, y=ylab, title=title)+
    theme_bw() + theme(panel.border = element_blank(),
                       # panel.grid.major = element_blank(),
                       # panel.grid.minor = element_blank(), 
                       axis.line = element_line(colour = "black") 
                        )
    
return(p)
}

#### NMB for symbicort cost ###
  sim_PSA_SYMB    <- f_wrapper(fxn=f_gen_psa_NMB,
                               n_sim=11,
                               trt="symbicort")  
  
  sim_PSA_no_trt  <- f_wrapper(fxn=f_gen_psa_NMB,
                               n_sim=11)
  
  
  sim_PSA_PRIM    <- f_wrapper(fxn=f_gen_psa_NMB,
                               n_sim=11,
                               trt = "primatene") 

  sim_PSA_SYMB$NMB = (100000*sim_PSA_SYMB$QALYs) - sim_PSA_SYMB$Costs
  sim_PSA_SYMB$c.sym <- seq(from=0, to=500, by=50)




### INMB ###
  sim_PSA_no_trt  <- f_wrapper(fxn=f_gen_psa_base,
                               n_sim=1) %>% slice(rep(1,16))
  sim_PSA_no_trt$WTP <- seq(from=0, to=150000, by=10000)
  sim_PSA_no_trt$NMB <- (sim_PSA_no_trt$WTP*sim_PSA_no_trt$QALYs) - sim_PSA_no_trt$Costs
  
  
  sim_PSA_PRIM    <- f_wrapper(fxn=f_gen_psa_base,
                               n_sim=1,
                               trt = "primatene") %>% slice(rep(1,16))
  sim_PSA_PRIM$WTP <- seq(from=0, to=150000, by=10000)
  sim_PSA_PRIM$NMB <- (sim_PSA_PRIM$WTP*sim_PSA_PRIM$QALYs) - sim_PSA_PRIM$Costs
  
#### INMB for symbicort compared to primatene and no trt
  sim_PSA_SYMB    <- f_wrapper(fxn=f_gen_psa_base,
                               n_sim=1,
                               trt="symbicort")  %>% slice(rep(1,16))
  sim_PSA_SYMB$WTP <- seq(from=0, to=150000, by=10000)
  sim_PSA_SYMB$NMB <- (sim_PSA_SYMB$WTP*sim_PSA_SYMB$QALYs) - sim_PSA_SYMB$Costs
  sim_PSA_SYMB$INMB_notrt <- sim_PSA_SYMB$NMB - sim_PSA_no_trt$NMB
  sim_PSA_SYMB$INMB_PRIM <- sim_PSA_SYMB$NMB - sim_PSA_PRIM$NMB

#plots
sym_PRIM<-NMB_plot(data=sim_PSA_SYMB,
         x=sim_PSA_SYMB$WTP,
         y=sim_PSA_SYMB$INMB_PRIM,
         xlab="Willingness to pay (WTP)",
         ylab="Incremental net monetary benefit (INMB)",
         title="Budesonide-fomoterol vs. Inhaled Epinephrine")
  
sym_notrt<-NMB_plot(data=sim_PSA_SYMB,
         x=sim_PSA_SYMB$WTP,
         y=sim_PSA_SYMB$INMB_notrt,
         xlab="Willingness to pay (WTP)",
         ylab="Incremental net monetary benefit (INMB)",
         title="Budesonide-fomoterol vs. No OTC Option")


theme_set(theme_pubr())
y <- ggarrange(sym_PRIM, sym_notrt,
               ncol=2,
               nrow=1,
               labels =c('A','B')) 

ggsave("INMB.png", plot=y, 
       path="C:\\Users\\Joseph\\Desktop\\Grad studies MSc and Applications\\AsthmaCEA2022\\Manuscript\\Figures",
       width= 11,
       height= 6,
       dpi = 1000)


```

### Extras ###
```{r, PSA parameter estimation}
### Estimating parameters
  estBetaParams <- function(mu, var) {
    alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
    beta <- alpha * (1 / mu - 1)
    return(params = list(alpha = alpha, beta = beta))
  }
  
  estGammaParams <- function(mu, var) {
    beta <- var/mu
    alpha <- mu/beta
    return(params = list(alpha = alpha, beta = beta))
  }
  
  
```
```{r, Base Case Analysis OLD}

sim_markov_BC <- function(){
##################################### Run the simulation ##################################
sim_markov_no_trt <- Markov(f_gen_psa_base(), v.M_1, n.t, v.n, d.c, d.e) # run for no treatment
sim_markov_PRIM    <- Markov(f_gen_psa_base(), v.M_1, n.t, v.n, d.c, d.e, trt = "primatene")  # run for treatement
sim_markov_SYMB    <- Markov(f_gen_psa_base(), v.M_1, n.t, v.n, d.c, d.e, trt = "symbicort")  # run for treatement

  
#storing vectors
  v.C <- c(sim_markov_no_trt$tc_hat, sim_markov_PRIM$tc_hat, sim_markov_SYMB$tc_hat) 
  v.E <- c(sim_markov_no_trt$te_hat, sim_markov_PRIM$te_hat, sim_markov_SYMB$te_hat)
  v.Cont <- c(sim_markov_no_trt$controlled, sim_markov_PRIM$controlled, sim_markov_SYMB$controlled)
  v.Exac <- c(sim_markov_no_trt$severe_exac, sim_markov_PRIM$severe_exac, sim_markov_SYMB$severe_exac)

# create the cost-effectiveness table
table_markov <- as_tibble(cbind(v.C, v.E, v.Cont, v.Exac))
table_markov$Strategy <- c("No treatment","Epinephrine","ICS-LABA")
table_markov <- select(table_markov, Strategy, everything())
colnames(table_markov) <- c("Strategy","Costs","QALYs", "Control","Exac")

# calculate ICER
table_markov <- table_markov %>%
                    arrange(desc(QALYs)) %>% # no treatment is dominated by symbicort 
                    mutate(delta_Costs = c(Costs[1]-Costs[3],Costs[2]-Costs[3], NA),  # calculate incremental costs
                           delta_QALYs = c(QALYs[1]-QALYs[3],QALYs[2]-QALYs[3],NA),  # calculate incremental QALYs
                           delta_control = c(Control[1]-Control[3],Control[2]-Control[3],NA),
                           delta_exac = c(Exac[1]-Exac[3],Exac[2]-Exac[3],NA),
                           ICER=delta_Costs/delta_QALYs) # calculate the ICER

return(table_markov)
}

# write.csv(table_markov,"CEA_markov_v1.csv")
#  write.csv(table_markov, "C:\\Users\\Joseph\\Desktop\\Grad studies MSc and Applications\\AsthmaCEA2022\\Preliminary results\\CEA_Markov_BC_v1.csv")
```
